Global code motion of congruent instructions

### 2017-04-18 paper.tex
remove irrelevant old text

\subsubsection{Legality of sinking expressions}
Sometimes, hoisting is not upward-safe \cite{click1995global}, e.g., if the
expressions are in a landing pad, sinking of those expressions
may reduce code size. Sinking may also reduce the register pressure in some
cases, e.g., when the use operands are not kills. For sinking, higher ranked
expressions would be sunk first. And it would be illegal to sink higher ranked
identical expressions if they are not anticipable in the common
post-dominator. For example:

@@

Overall there was only a minor change in performance regressed with a few ups
(429.mcf, 458.sjeng etc.) and a few regresssions (447.dealII, 453.povray
etc.). The code-motion pass was run twice in the pass pipeline (at -Ofast),
first after EarlyCSE pass and then after the inliner. Because EarlyCSE removes
local redundancies, GCM would have to analyze less redundant instructions, also
as inliner creates more redundancies, GCM would have more opportunities to move
congruent instructions out of branches.

Code size also remained almost the same as a result of code-motion as shown in
Table~\ref{tab:code-size}.  The code size listed here is the text size (from
linux size command) of the final executables for each benchmark.  Some bencmarks
like 473.astar increased in code size by almost 2\%. This is because once the
code-size of a function decreases it becomes cheaper to inline. As we can see in
Table-\ref{tab:stats}, four more functions got inlined in 473.astar. Except for
447.dealII, 450.soplex, 482.sphinx3 and 483.xalancbmk all other benchmarks got
more functions inlined.

The results for code-hoisting are listed in
Table~\ref{tab:code-motion-metric}. The table lists the number of scalars,
loads, stores and calls hoisted as well as removed. For each category, the
number of instructions removed is greater or equal to the number of instructions
hoisted because each hoisting is performed only when at least one identical
computation is found.

Loads are hoisted the most followed by scalars, stores and calls in decreasing
order.  This was the common trend in all our experiments. One reason why loads
are hoisted the most is the early execution of this pass (before mem2reg) in the
\LLVM{} pass pipeline. Passes like mem2reg, instcombine might actually remove
those loads so this order may change should this pass be scheduled later. The number
of instructions sunk is very less because \LLVM{} has a code sinking pass inside SimplifyCFG
which is run before the global code-motion and sinks some instructions. Another reason is
that instructions for which dependencies are not directly available in the successor is
not sunk for now. Sinking those instructions would require a look ahead on the sinkability
of the dependencies. We plan to implement this in near future.

Other static metrics are listed in Table~\ref{tab:stats} and in
Table~\ref{tab:code-motion-metric}.  Here we can see that except for
rematerializing defs for splitting, which has an overhead of ~2\%, all other
parameters have less than 1\% overhead. This is to explain why the performance
does not go down with our implementation (and cost-model) of code hoisting pass.

>>>>

\subsubsection{Legality of sinking expressions}
\label{subsec:legality-sink}
Sometimes, hoisting is not upward-safe \cite{click1995global}, e.g., if the
expressions are in a landing pad, sinking of those expressions
may reduce code size. Sinking may also reduce the register pressure in some
cases, e.g., when the use operands are not kills. For sinking, higher ranked
expressions would be sunk first. And it would be illegal to sink higher ranked
identical expressions if they are not anticipable in the common
post-dominator. For example:

@@

Overall there was only a minor change in performance regressed with a few ups
(429.mcf, 458.sjeng etc.) and a few regresssions (447.dealII, 453.povray
etc.). The code-motion pass was run twice in the pass pipeline (at -Ofast),
first after EarlyCSE pass, and then after the inliner. Because EarlyCSE removes
local redundancies, GCM would have to analyze less redundant instructions, also
as inliner creates more redundancies, GCM would have more opportunities to move
congruent instructions out of branches.

Code size also remained almost the same as a result of code-motion as shown in
Table~\ref{tab:code-size}.  The code size listed here is the text size (from
linux size command) of the final executables for each benchmark.  Some bencmarks
like 473.astar increased in code size by almost 2\%. This is because once the
code-size of a function decreases (due to GCM), it becomes cheaper to inline. As
we can see in Table-\ref{tab:stats}, four more functions got inlined in
473.astar. Except for 447.dealII, 450.soplex, 482.sphinx3 and 483.xalancbmk all
other benchmarks got more functions inlined.

The compile time metrics of GCM are listed in
Table~\ref{tab:code-motion-metric}. The table lists the number of scalars,
loads, stores and calls hoisted as well as removed. For each category, the
number of instructions removed is greater or equal to the number of instructions
hoisted because each hoisting is performed only when at least one identical
computation is found. Loads are hoisted the most followed by scalars, stores and
calls in decreasing order.  This was the common trend in all our
experiments. One reason why loads are hoisted the most is the early execution of
this pass (before mem2reg) in the \LLVM{} pass pipeline. Passes like mem2reg,
instcombine might actually remove those loads so this order may change should
this pass be scheduled later. The number of instructions sunk is very less
because \LLVM{} has a code sinking pass inside SimplifyCFG which runs before the
GCM and sinks some instructions. Another reason is that instructions for which
dependencies are not directly available in the successor
(Section~\ref{subsec:legality-sink}) is not sunk for now. Sinking those
instructions would require a look ahead on the sinkability of the
dependencies. We plan to implement this in near future.
